---
title: "Topic-4"
author: "TA Kiran"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Set the Working Directory:
setwd("/Users/kirankumarisheshma/Desktop/Introduction To data science/Topic-4")

#Get Current Working Directory
getwd()
```

**scatter plot **
-Scatter plots are particularly useful for visualizing the relationship or correlation between two continuous variables

-Use the plot() function to create a scatter plot.
```{r}
# Load the iris dataset
data("iris")


# Create the scatterplot and add labels
plot(iris$Sepal.Length, iris$Sepal.Width, 
     main = "Scatterplot Example", xlab = "Sepal.Length", ylab = "Sepal.Width", pch = 19, col="blue")
```


**Label Data Points in the Scatter plot**  

-You can use the text() function to add labels or annotations to specific data points on the scatter plot.**
```{r}
# Create a scatter plot
plot(iris$Sepal.Length, iris$Sepal.Width, main = "Sepal Length vs. Sepal Width", xlab = "Sepal Length", ylab = "Sepal Width", pch = 19, col = "blue")

# Add labels for the first 100 data points
text(iris$Sepal.Length[1:100], iris$Sepal.Width[1:100], labels = iris$Species[1:100], cex = 0.75, font = 2) #cex =character expansion
 
```

-You create a legend using the legend() function, specifying its position, labels, colors, point characters, and title.
```{r}
##### scatter plot
plot(iris$Sepal.Length, iris$Sepal.Width, pch=19, main='Scatter plot',
     xlab='Speal.Length', ylab = 'Sepal.Width', cex=0.75,
     col=iris$Species)
legend("topright", legend = unique(iris$Species),col=unique(iris$Species), pch = 19,)
 
```

**Points and Lines in the Scatter plot  **
-In a scatter plot, you can control the appearance of points and lines to customize the visualization according to your needs. 
```{r}
s1 = which(iris$Species == "setosa") 
plot(iris$Sepal.Length, iris$Sepal.Width, main = "Scatterplot Example", xlab = "Sepal.Length", ylab = "Sepal.Width", type = "n")
points(iris$Sepal.Length[s1], iris$Sepal.Width[s1], pch = 2,col="red")

```

```{r}
s1 = which(iris$Species == "setosa")
s2 = which(iris$Species == "versicolor")

# Create an empty scatter plot canvas
plot(iris$Sepal.Length, iris$Sepal.Width, main = "Scatterplot Example", xlab = "Sepal.Length", ylab = "Sepal.Width", type = "n")

# Plot data points for "setosa" with filled triangles (pch=2) and color red (col=2)
points(iris$Sepal.Length[s1], iris$Sepal.Width[s1], pch = 2, col = "red")

# Plot data points for "versicolor" with filled squares (pch=3) and color green (col=3)
points(iris$Sepal.Length[s2], iris$Sepal.Width[s2], pch = 3, col = "blue")

# Add dashed lines connecting two points
lines(c(4.5, 6.5), c(2, 4.5), lty = 2)

# Create a legend to explain the species and their colors
legend(x = 7, y = 2.8, c("setosa", "versicolor"), col = c("red", "blue"), pch = c(2, 3), title = "Species")

```

**Histogram**

-A histogram is a graphical representation of the distribution of a dataset. It is a type of data visualization that displays the frequency or count of data points within predefined intervals or bins

-  "bins" and "breaks"  both refer to the intervals or divisions used to group data points in a visualization that displays the distribution of data.

-breaks = 5 means you are dividing the data into 5 equally spaced bins. 

```{r}
##### histogram
hist(iris$Petal.Length, main='Histogram of Petal.Length', xlab='Petal.Length', col='skyblue3',
     breaks=5) # try breaks = 15 or a vector
```

**Bar plot**

-Bar plots are typically used to visualize and represent categorical or discrete data

```{r}
##### bar plot
barplot(table(iris$Species), main='Species',col=c("red","blue","green")) #table(iris$Species)It counts the number of occurrences of each unique species (setosa, versicolor, and virginica).
legend("topright", legend = c("setosa", "versicolor", "virginica"),fill = c("red", "blue", "green"))

```

**Box plot**

-Box plots are primarily used for visualizing and summarizing the distribution of numerical data, specifically continuous or interval data. 

```{r}
##### box plot
boxplot(iris[,-5], main='Boxplot',col="red")
boxplot(Petal.Length~Species, data=iris)
```

- from the above box plot we can see that the data are Skewed Data. there are outliers in the data set.

**Density plot**

```{r}
# Load the ggplot2 library for creating density plots
library(ggplot2)

# Create a density plot for Sepal.Length
ggplot(iris, aes(x = Sepal.Length, fill = Species)) +
  geom_density(alpha = 0.5) +
  labs(
    title = "Density Plot of Sepal.Length by Species",
    x = "Sepal Length",
    y = "Density"
  )

```

```{r}
x <- rnorm(1000,0,1)
kde <- density(x)
hist(x, ylim=c(0,max(kde$y)), freq=F, col=rainbow(24))
lines(kde, col='blue', lwd=5)
```

**QQ-Plot¶**

- qqnorm() creates the QQ plot, and qqline() adds a reference line to the plot. If the points in the plot closely follow the reference line, it suggests that the data is approximately normally distributed.

```{r}
# Generate example data
data <- rnorm(100)  # Generating normally distributed data

# Create a QQ plot against the normal distribution
qqnorm(data)
qqline(data)
```

**##### pie chart**
```{r}
pie(table(iris$Species))
```

```{r}

# Calculate the counts of each unique species
species_counts <- table(iris$Species)

# Define custom colors
colors <- c("red", "blue", "green")

# Create a pie chart with percentages and custom colors
pie(species_counts, labels = paste(names(species_counts), "(", round(100 * species_counts / sum(species_counts), 1), "%)"), col = colors)

# Set plot title
title("Pie Chart of Species Distribution")

# Add a legend
legend("topright", legend = names(species_counts), fill = colors)

```


**Curve**

```{r}
##### curve
curve(x**2+sin(x)+exp(x)*cos(x), 0,5)
```

**3D Density**
```{r}
grid_x <- seq(-10,10,length=100)
grid_y <- grid_x
height <- matrix(0, nrow=length(grid_x), ncol=length(grid_y))
for(i in 1:length(grid_x)){
  for(j in 1:length(grid_y)){
    height[i,j] <- grid_x[i]**2 + grid_y[j]**2
  }
}
persp(grid_x, grid_y, height)
contour(grid_x, grid_y, height)
image(grid_x, grid_y, height)
```

**Building ggplot2**

```{r}
# Load dplyr package explicitly for the pipe operator
library(dplyr)

# Now you can use %>% for data manipulation
iris %>%
  filter(Sepal.Length > 5) %>%
  select(Sepal.Length, Sepal.Width)

# Load ggplot2 separately if needed
library(ggplot2)

# Now you can use ggplot() for data visualization
iris %>%
  ggplot() +
  geom_point(aes(x = Sepal.Length, y = Sepal.Width))

```

**1.Aesthetics : Specify the mapping manner**

- aes(x, y, color, size, shape, alpha): where

- x: x axis data

- y: y axis data

- color: color function

- size: geometric attribute size

- shape: geometric attribute shape

- alpha: transparency

```{r}
iris %>% ggplot() + 
  aes(x=Petal.Length, y=Petal.Width, color=Species)
```

**2.Geometries : Specify the geometric attribute**
```{r}
iris %>% ggplot() + 
  aes(x=Petal.Length, y=Petal.Width, color=Species) + 
  geom_point()
```

**Other commonly used function**

- ggtitle():set title

- xlab():set x axis title

- ylab():set y axis title

Or we can just simply use the following function

- labs(x,y,title,fill)

**geom_boxplot**

```{r}
ggplot(iris) + aes(x=Species, y=Petal.Length, fill=Species) + 
  geom_boxplot() + labs(x='Species', y='PetalLength', 
                        fill='Species~~',title=' Box plot')
```

**ggplot2 Geometries**

*geom_bar*

- geom_bar(position)
- dodge: dodge bars
- fill: fill bars
- stack: stack bars

```{r}
ggplot(mpg) + aes(x=fl,fill=drv) + 
  geom_bar(position='dodge') + # change to fill or stack
  coord_flip() + theme_minimal() + 
  scale_fill_brewer(palette="Blues")
```

**geom_point**

```{r}
ggplot(mtcars) + aes(x=disp, y=mpg, col=as.factor(gear),size=5) + 
  geom_point(position='jitter') + theme_minimal() + 
  scale_color_brewer()
```

**geom_line**
```{r}
library(gapminder)

gapminder %>% filter(country %in% c('China', 'Taiwan')) %>% 
ggplot() + aes(x = year, y = gdpPercap, colour = country) + 
  geom_line() + theme_minimal()
```

**geom_boxplot**
```{r}
ggplot(gapminder, aes(x = continent, y = gdpPercap)) +
  geom_boxplot()
```

**geom_histogram**
-stat_bin()` using `bins = 30`. Pick better value with `binwidth`.
```{r}
ggplot(iris) + aes(x=Petal.Length) + stat_bin(bins=30) + 
  geom_histogram()
```






