---
title: "Topic_3 Introduction to Data Science"
author: "TA Kiran"
date: "2023-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Set the Working Directory:
setwd("/Users/kirankumarisheshma/Desktop/Introduction To data science/Topic-3")

#Get Current Working Directory
getwd()
```

**Reading Data – Case Study: Health Survey**
```{r}
data <- read.csv("survey.csv") #read csv file
## View the first few rows of the data frame
head(data, 10)

## View the last few rows of the data frame
tail(data,5)

```

```{r}
#class() function is used to determine the class or data type of an object
class(data)
# Check the structure of the data frame
str(data)
# View summary statistics of the data frame
summary(data)
```

```{r}
data[c(1, 2, 3, 4, 5,6,7), ]
data[1:10, ]
```

**convert factor into ordinal variable**

```{r}
# Load the 'ordinal' package
data$smoke
```

```{r}
# Load the 'ordinal' package
library(ordinal)

data$smoke <- ordered(data$smoke,
                      levels = c("Never", "Occas", "Regul", "Heavy"))
data$smoke
```


```{r}
#Another Example of the Ordinal() Function Use

# Create an ordinal variable
satisfaction <- ordered(c("Very Dissatisfied", "Dissatisfied", "Neutral", "Satisfied", "Very Satisfied"), 
                        levels = c("Very Dissatisfied", "Dissatisfied", "Neutral", "Satisfied", "Very Satisfied"))

# Print the ordinal variable
print(satisfaction)
```

**unique() function**

-  It is a helpful function for identifying distinct values within a dataset. 
```{r}
##Cleaning up data - Data entry error
unique(data$sex) #get unique value
unique(data$height)
unique(data[2])  # using Index
```

**Use the For loop to get all togther**
```{r}
# Create an empty list to store unique values for each column
unique_values <- list()

# Loop through columns using a for loop
for (col_name in colnames(data)) {
  unique_values <- unique(data[col_name])
  #print(unique_values) 
  # Print unique values for each column
  for (col_name in names(unique_values)) {
  cat("Unique values in column", col_name, ":", unique_values[[col_name]], "\n")
}#Access the column by name
}

```

**using the lapply () function**
```{r}
unique_values <- lapply(data, unique)
print(unique_values)
```

**Missing Values**
- In R, the complete.cases() function is used to identify rows in a data frame or matrix that have complete cases, i.e., rows with no missing (NA) values.
```{r}
# Count rows with complete cases (no missing values)
cat("Number of complete rows:", sum(complete.cases(data)), "\n")

# Count rows with missing values
cat("Number of rows with missing values:", sum(!complete.cases(data)), "\n") 
```
**Check for the total number of missing values in each column**

```{r}
# Count rows with missing values
cat("Number of rows with missing values:", sum(!complete.cases(data$sex)), "\n") 
# Count rows with missing values
cat("Number of rows with missing values:", sum(!complete.cases(data$height)), "\n") 
# Count rows with missing values
cat("Number of rows with missing values:", sum(!complete.cases(data$weight)), "\n") 
```
**Check for the total number of missing values in each column in a data frame**
```{r}
# Create an empty list to store missing values for each column
missing_values <- list()

# Loop through columns using a for loop
for (col_name in colnames(data)) {
  missing_values <- sum(!complete.cases(data[col_name]))
  cat("missing values in column", col_name, ":", missing_values, "\n")
}
```
**Check for the Position of the Missing values in each columns**

- In R, you can use the which(is.na()) combination of functions to identify the indices (positions) of missing (NA) values within a vector, matrix, or data frame.
```{r}

###Cleaning up data - Missing value
which(is.na(data$sex)) #check where is the missing value
which(is.na(data$height)) #check where is the missing value
```

```{r}
# Create an empty list to store missing values position for each column
position_of_Missing_values <- list()

# Loop through columns using a for loop
for (col_name in colnames(data)) {
  position_of_Missing_values <- which(is.na(data[col_name])) 
  cat("position of the missing values in each column", col_name, ":", position_of_Missing_values, "\n")
}
```

**Cleaning the dataset**
```{r}
data$sex[data$sex == 'F'] = 'Female' # replace 'F" to 'Female'
data$sex[data$sex == 'M'] = 'Male'   # replace 'M' to 'Male'
unique(data$sex)
```

**Replace continuous data with mean**

```{r}
#Calculate the mean of female heights (ignoring missing values)
female.height <- mean(data$height[which(data$sex == 'Female')], na.rm=T)
female.height
## Calculate the mean of male heights (ignoring missing values)
male.height <- mean(data$height[which(data$sex == 'Male')], na.rm=T)
male.height
#replace NA with mean of female & male height 
data$height[which(data$sex == 'Female' & is.na(data$height))] = female.height
data$height[which(data$sex == 'Male' & is.na(data$height))] = male.height
unique(data$height)
```
**Replace categorical data with bootstrap in R**

```{r}
bootstrap <- function(series){
  ## Filter the series to keep only complete cases
    tb <- table(series[complete.cases(series)])
    ## Calculate the probability distribution based on frequencies
    prob <- tb / sum(tb)
    ## Generate a sample based on the probability distribution
    smpl <- sample(names(tb), prob=prob, size=length(sum(!complete.cases(series))))
    ## Replace missing values in the original series with the sample
    series[!complete.cases(series)] <- smpl
    return (series)
}
```

```{r}
#data$smoke <- ordered(data$smoke, levels=c('Never', 'Occas', 'Regul', 'Heavy'))
#data$smoke[!complete.cases(data)] <- names(which.max(table(data$smoke)))

#data$exercise <- ordered(data$smoke, levels=c('None', 'Some', 'Freq'))
#data$exercise[!complete.cases(data)] <- names(which.max(table(data$exercise)))

data$smoke <- bootstrap(data$smoke)
data$handedness <- bootstrap(data$handedness)
data$sex <- bootstrap(data$sex)

cat("Numberof non NA row:", dim(data[complete.cases(data),])[1])
cat("\nNumberof non NA row:", dim(data[!complete.cases(data),])[1])

summary(data)
```




